---
title: "Activity 5 - Jawadul Chowdhury"
date: 'Due Date: April 18, 2025'
output: html_notebook
---

## Problem 1:
Contained in the file "Activity5PGA.csv" (found in Canvas) is data for the average driving distance for each PGA 
Tour golf from the 2022-2023 season. Make sure that file is in the same folder as this .Rmd file, and run the chunk 
of code below so that you can use it. I've called the dataset PGA, but you can certainly rename if it you'd like.

```{r}
# creating a PGA dataframe
pga_df <- read.csv("Activity5PGA.csv")
```

Your goal for this problem is to use a random sample of size 25 to create confidence intervals for the mean and 
standard deviation.

### Part (a)
First, we need a random sample of size 25. Use the set.seed() function first, so that your specific results are 
reproducible. Then use the sample() function to create a random sample of size 25 from the "AVG" column of the data. 
I would recommend assigning your sample a name (such as "mysample") for ease of use in future parts.

```{r}
# setting the seed to allow for reproducible results
set.seed(42)

# creating a random sample of size 25
sample1 <- sample(pga_df$AVG, 25)
print(sample1)

```


### Part (b)
Since we are in the "small sample" case, we need to assume our population distribution is (at least approximately) 
normal. One way to see if this is reasonable is by using a quantile-quantile (Q-Q) plot, which plots theoretical 
quantiles on the x-axis and sample quantiles on the y-axis. 

If these points fairly closely follow a straight diagonal line, the normality assumption is considered reasonable. 
Good news: R has built-in functions for such plots: qqnorm() will create the Q-Q plot, and qqline() will overlay the 
diagonal line. Use these functions to create the plot and line for your sample. 

For this problem, don't worry too much about if your plot doesn't seem to follow the line very  closely - I just want 
you to know how to create such a plot for now.

```{r}

# checking for presence of missing values
missing_col <- names(pga_df)[colSums(is.na(pga_df)) == nrow(pga_df)]
print(missing_col)

# since there are no missing values, we create the qq plot
qqnorm()


```


### Part (c)
Based on your sample, create a 95% (two-sided) confidence interval for the average driving distance for PGA Tour players during the 2022-2023 season. Use the code chunk below for any calculations, and give your interval as well as an interpretation sentence where indicated. Be sure to be appropriately descriptive/specific/clear with your interpretation (note that driving distance here are measured in yards).

```{r}

```

**Answers for part (c): **

### Part (d)
If you repeated this process 50 times, how many of those 50 confidence intervals would you expect to contain the true mean? Does the interval you created contain the true mean? Note that we would not usually know this, but here we can check just for fun :) 
(Type your answers where below as indicated.)

**Answers for part (d):**

### Part (e)
Based on your sample, create a 95% upper confidence bound for the standard deviation of driving distance for PGA Tour players during the 2022-2023 season. Use the code chunk below for any calculations, and give your interval as well as an interpretation sentence where indicated. Be sure to be appropriately descriptive/specific/clear with your interpretation.

```{r}

```

**Answers for part (e): **


## Problem 2
Here you will be creating a 90% confidence interval for a proportion based on data you collected. For each part, type your answers in the appropriate places.

### Part (a)
What is the "topic" you chose? In other words, what exactly are you creating a confidence interval for?

**Answer for part (a):**

### Part (b)
Briefly describe how your data was collected, and what the results were. (Be sure to include things like sample size and sample proportion.)

**Answer for part (b):**

### Part (c)
State your confidence interval, along with an appropriately specific/descriptive/clear interpretation sentence. You may use the code chunk below for any calculations (or you may use a calculator and z-table instead if you prefer - just make sure to indicate this if that is your choice).

```{r}
#Use this code chunk for any calculation for Problem 2
```

**Answer for part (c):**


## Problem 3:

According to the city of Tulsa's municipal website, the rate it charges per 5 CCF of residential was $22.62 in 2020. We are going to investigate how the residential water rates of other US public utilities compare to Tulsa's rate. To do this, we will conduct a hypothesis test, with significance level 0.05 to try to determine whether the population average rate of 5 CCF of residential water in the US in 2020 was less than the rate charged by Tulsa.

The file “Activity5Water.csv'” (found in Canvas) contains the rate per 5 CCF of residential water for 42 US cities. If you make sure that the file is in the same folder as this file, you should be able to use it with no problems as long as you run the following chunk of code. I've just called it "water" for ease of use, but you can of course change that if you want to.

```{r}
water <- read.csv("Activity5Water.csv")
```

Include any calculations used in parts (a)-(e) in the code chunk below, and give an answer/response for each of the parts in the appropriate space.

```{r}
#Use this chunk of code for any calculations in parts(a)-(e)

```

### Part (a) 
First, determine the null and alternative hypotheses.

**Answer for part (a):**

### Part (b)
Next, calculate our test statistic value.

**Answer for part (b):**

### Part (c)
Now use the value from part (b) to find the P-value.

**Answer for part (c):**

### Part (d) 
Is the null hypothesis rejected at our chosen significance level?

**Answer for part (d):**

### Part (e)
Give an interpretation of the result in the context of this problem.

**Answer for (e):**

### Part (f) 
In practice, even with a large sample size, a t-test is actually preferred (instead of a z-test) if the tools to do so are available and the normality assumption is reasonable. In the R Console, type in "?t.test" and hit enter to view the documentation for the t.test() function. Then, in the code chunk below, use this function to "redo" the hypothesis test we just did (except now we are using the t-distribution) with just a single line of code. Make sure you specify the appropriate arguments in the t.test() function.

```{r}
#This code just is just for using the t.test() function for part (f)

```
